---
title: "p8105_hw3_kb3066"
author: "Kaila Boyer"
date: "10/14/2021"
output: github_document
---

```{r}
library(tidyverse)
library(p8105.datasets)
library(patchwork)
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  fig.width = 7,
  fig.asp = .7,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis",
  scale_colour_discrete = scale_colour_viridis_d,
scale_fill_discrete = scale_fill_viridis_d
)


```
## Problem 1

The goal is to do some exploration of this dataset. To that end, write a short description of the dataset, noting the size and structure of the data, describing some key variables, and giving illstrative examples of observations. Then, do or answer the following (commenting on the results of each):

How many aisles are there, and which aisles are the most items ordered from?

```{r}
data("instacart")

instacart %>% summarize(n_aisles = n_distinct(aisle_id))

instacart %>% select(aisle_id, product_id, aisle) %>% 
  group_by(aisle, aisle_id) %>% count() %>% 
  arrange(desc(n)) 

  instacart %>%  
    summarise( avg_prod_num = mean(order_number, na.rm = T), 
               total_users = n_distinct(user_id), 
               deptartments = n_distinct(department_id)) 

  instacart %>%  
  ggplot(aes(x = department, product_id)) + 
  geom_col() + 
  labs( 
      y = "Number of products", 
      x = "Department") + 
  theme(axis.text.x = element_text(angle = 90)) 
```
There are 134 unique aisles. The aisles with the most items ordered are 83 and 24 with over 150,000 items (fresh vegetables and fresh fruits). 

There average number of products in an order is 17.1, there are 21 departments, and 131,209 users in the dataset. 

The bar graph shows the number of products in each department. 


Make a plot that shows the number of items ordered in each aisle, limiting this to aisles with more than 10000 items ordered. Arrange aisles sensibly, and organize your plot so others can read it.
```{r }
instacart %>% 
  group_by(aisle) %>% count() %>% 
  arrange(desc(n)) %>% 
  filter(n >10000) %>% 
  transform(aisle = reorder(aisle, order(n, decreasing =F))) %>% 
  ggplot(aes(x = aisle, y = n)) + 
  geom_col(fill = "dark green") + 
  coord_flip() + 
  labs(
    title = "Number of Items in Aisles", 
    x = "Aisle Name", 
    y= "Count"
  )
```
The aisles for fresh fruit and fresh vegetables have the most items. 


Make a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”. Include the number of times each item is ordered in your table.
```{r}
instacart %>% select(aisle, product_name) %>% 
  group_by(aisle) %>% 
  filter(aisle == "baking ingredients" | aisle == "dog food care" | aisle == "packaged vegetables fruits") %>%  
  count(product_name) %>%
  slice_max(order_by = n, n = 3) %>% 
  knitr::kable()
```
This table shows the top three items in each of the three aisles, with the number of times each products is ordered in the column titled "n."


Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers (i.e. produce a 2 x 7 table).
```{r error=FALSE, message=FALSE, warning=FALSE}

instacart %>% filter(product_name == "Pink Lady Apples" | product_name == "Coffee Ice Cream") %>%
  select(product_name, order_dow, order_hour_of_day) %>%
  group_by(order_dow, product_name) %>%  
  mutate(order_dow = if_else(order_dow == 0, "Sunday", 
                     if_else(order_dow==1, "Monday", 
                     if_else(order_dow==2, "Tuesday", 
                     if_else(order_dow==3, "Wednesday", 
                     if_else(order_dow==4, "Thursday", 
                     if_else(order_dow==5, "Friday", 
                     if_else(order_dow==6, "Saturday", NA_character_)))))))) %>%  
  summarize(mean_hr = mean(order_hour_of_day, na.rm = T)) %>%  
  pivot_wider(
    names_from = order_dow, 
    values_from = mean_hr
  ) %>% relocate(product_name, Sunday, Monday, Tuesday, Wednesday, Thursday, Friday, Saturday) %>% 
  knitr::kable(caption = "Mean time of order each day of week")
```
The mean time of ordering Pink Lady apples and coffee ice cream for each day of the week, in military time. The mean time of ordering is mostly in the late morning and early afternoon. 


## Problem 2

This problem uses the BRFSS data. DO NOT include this dataset in your local data directory; instead, load the data from the p8105.datasets package.

First, do some data cleaning:

format the data to use appropriate variable names;
focus on the “Overall Health” topic
include only responses from “Excellent” to “Poor”
organize responses as a factor taking levels ordered from “Poor” to “Excellent”
Using this dataset, do or answer the following (commenting on the results of each):


```{r}
data("brfss_smart2010")

brfss_smart2010 <- brfss_smart2010 %>% janitor::clean_names() %>% 
  filter(topic == "Overall Health") %>% 
  filter(response == "Poor" | response =="Fair" | response == "Good" | response == "Very good" | response== "Excellent") %>% 
  mutate(response = factor(response, levels = c("Poor", "Fair", "Good", "Very good", "Excellent")))
```

In 2002, which states were observed at 7 or more locations? What about in 2010?

```{r}
brfss_smart2010 %>% filter(year == 2002) %>% group_by(locationabbr) %>% distinct(locationdesc) %>% count() %>% filter(n >=7)

brfss_smart2010 %>% filter(year == 2010) %>% group_by(locationabbr) %>% distinct(locationdesc) %>% count() %>% filter(n >=7)
```
In 2002, Connecticut, Florida, Massachusetts, North Carolina, New Jersey, and Pennsylvania had 7 or more locations. 

In 2010, California, Colorado, Florida, Massachusetts, Maryland, North Carolina, Nebraska, New Jersey, New York, Ohio, Pennsylvania, South Carolina, Texas, and Washington had 7 or more locations. 


Construct a dataset that is limited to Excellent responses, and contains, year, state, and a variable that averages the data_value across locations within a state. Make a “spaghetti” plot of this average value over time within a state (that is, make a plot showing a line for each state across years – the geom_line geometry and group aesthetic will help).
```{r}
prob2 <- brfss_smart2010 %>% 
  filter(response == "Excellent") %>% 
  select(year,locationdesc, data_value ) %>% 
  group_by(locationdesc ) %>%
  mutate(data_avg = mean(data_value, na.rm = T)) 

brfss_smart2010 %>% 
  filter(response == "Excellent") %>% 
  select(year, locationabbr, data_value ) %>% 
  group_by(year, locationabbr) %>%
  mutate(data_avg = mean(data_value, na.rm = T)) %>% 
  ggplot(aes(y = data_avg, x = year)) + 
  geom_smooth(aes(color= locationabbr), se = FALSE) + 
  scale_color_discrete(name = "State") +
  xlab("Year") +  
  ylab("Average Data Value") + 
  ggtitle("Average values among 'Excellent' responses in each state") +
  theme(legend.position = "right"
        ) 
```


Make a two-panel plot showing, for the years 2006, and 2010, distribution of data_value for responses (“Poor” to “Excellent”) among locations in NY State.
```{r}
ny <- brfss_smart2010 %>% filter(locationabbr == "NY")  %>% select(response, data_value, year) 

ny06 <- ny %>% filter(year == 2006) %>% 
  ggplot(aes(y = data_value, x = year, fill = response )) + 
  geom_col(position = "dodge") + 
  scale_x_continuous( breaks = c( )) + 
  ylim(0, 40) +
  labs(
    x = "",  
    y = "Data Value", 
    fill = "Responses") + 
  ggtitle("2006") + 
  theme(plot.title = element_text(hjust = 0.5)) 

ny10 <- ny %>% filter(year == 2010) %>% 
  ggplot(aes(y = data_value, x = year, fill = response )) + 
  geom_col(position = "dodge") + 
  scale_x_continuous( breaks = c(  )) + 
  labs(
    x = "",  
    y = "  ") + 
  ggtitle("2010") + 
  theme(legend.position = "none", 
        plot.title = element_text(hjust = 0.5))  
  
ny06 + ny10 + plot_layout(guides = "collect") + 
  plot_annotation(title = "Distribution of Data Values by Response in NY", 
  theme = theme(plot.title = element_text(size = 16, hjust= 0.6)) )
```
In 2006, the most data fell into "Good" and in 2010 most of the data fell into "Very good."


## Problem 3 
Accelerometers have become an appealing alternative to self-report techniques for studying physical activity in observational studies and clinical trials, largely because of their relative objectivity. During observation periods, the devices measure “activity counts” in a short period; one-minute intervals are common. Because accelerometers can be worn comfortably and unobtrusively, they produce around-the-clock observations.

This problem uses five weeks of accelerometer data collected on a 63 year-old male with BMI 25, who was admitted to the Advanced Cardiac Care Center of Columbia University Medical Center and diagnosed with congestive heart failure (CHF). The data can be downloaded here. In this spreadsheet, variables activity.* are the activity counts for each minute of a 24-hour day starting at midnight.

Load, tidy, and otherwise wrangle the data. Your final dataset should include all originally observed variables and values; have useful variable names; include a weekday vs weekend variable; and encode data with reasonable variable classes. Describe the resulting dataset (e.g. what variables exist, how many observations, etc).

```{r}
accel <- read_csv("accel_data.csv") %>% janitor::clean_names() %>% 
  mutate(day_type = if_else(day == "Saturday" | day == "Sunday", "Weekend", "Weekday")) %>% 
  select(week, day_id, day, day_type, everything())
dim(accel)
```
In the accel dataset, there are 35 rows, one for each day of the week for 5 weeks. There are a total of 1444 columns, where 1440 have activity data for each minute of the day. Additional variables are the week number, the day id, the day of the week and whether the day was a weekday or weekend.  


Traditional analyses of accelerometer data focus on the total activity over the day. Using your tidied dataset, aggregate accross minutes to create a total activity variable for each day, and create a table showing these totals. Are any trends apparent?

```{r}
accel <- accel %>% mutate(day_total = rowSums(across(activity_1:activity_1440, na.rm = T)))

accel  %>% group_by(day) %>% select(week, day, day_total) %>% 
  pivot_wider( names_from = day, 
               values_from = day_total) %>% 
  select(week, Sunday, Monday, Tuesday, Wednesday, Thursday, Friday, Saturday) %>% 
  knitr::kable()
```
The activity throughout the weeks generally increases from Monday to Sunday, there are no clear trends. Many of the days have a total activity ranging from 300,000 to 500,000. 


Accelerometer data allows the inspection activity over the course of the day. Make a single-panel plot that shows the 24-hour activity time courses for each day and use color to indicate day of the week. Describe in words any patterns or conclusions you can make based on this graph.
```{r}
prob3 <- accel %>%
  pivot_longer(activity_1:activity_1440,
    names_to = "activity",
    values_to = "value"
  ) %>% 
  mutate(activity = as.factor(activity))
  
prob3 %>% 
  ggplot(aes(x = activity, y=value)) +
  geom_line(aes(color = day), alpha = 0.5)  +
  xlab("Activity") + 
  ylab("Value") + 
  ggtitle("Activity over 24 hours")+ 
  theme(legend.position = "bottom", 
        plot.title = element_text(hjust = 0.5), 
        legend.text = element_text(size = 10)) + 
  scale_color_discrete(name = "Day")  + 
  scale_x_discrete(breaks = c(prob3[1 , 6], prob3[25200 , 6], prob3[ 50400, 6]), 
                     labels = c("12:01 am", "12 pm", "12 am")) + 
  theme(axis.text.x = element_text(angle = 90))
```
Appear to more active in Monday morning (orange color) and more active Sunday afternoon/evening (turquoise color). There is a dip in activity for all days towards mid day. 
